Make any stupid plot you want

### 2021-05-10 draft/sections/intro.tex
intro read through

Visualizations, by definition, reflect something of the underlying structure and semantics\cite{friendlyBriefHistoryData2008} of the data, whether through direct mappings from data into visual elements or via figurative representations that have meaning due to their similarity in shape to external concepts \cite{byrneAcquiredCodesMeaning2016}. We define visualization components to be structure preserving maps from data to visual representations. As illustrated in \autoref{fig:intro:artist}, visualizations map weather station precipitation data (a) into graphical elements such as points (b) and lines (c) and map individual data components (columns) into components of the visualization such as position or color. For example, the gray position encoder function converts the latitude and longitude to x and y positions, and the color encoders map temperature and station name to colors. A \texttt{point} function composites these encodings into attributes of a point in the map (a). The \texttt{line} function composites encoders that convert station name to color and date and precipitation to x and y positions into a piece of a line in the time series plot (c). We identify the structure these maps must preserve as the \textit{continuity} of the data and the \textit{equivariance} of the data and visual components.

We introduce a model of visualization components based on these \textit{continuity} and \textit{equivariance} constraints and use this model to develop a design specification. Our specification is targeted at building block libraries, as defined by Wongsuphasawat\cite{wongsuphasawatNavigatingWideWorld2021}, because they are the visualization tools that provide independent functions that map components of data to visual representations.  Just as the number of ways to arrange physical building blocks is solely constrained by the size and shape of the blocks, we propose that the only restriction on how building block library components can be composed are that the compositions preserve \textit{continuity} and \textit{equivariance}. Independent, modular, components are inherently functional\cite{hughesWhyFunctionalProgramming1989}, so we propose a functional architecture for our model. Doing so means the functional architecture can be evaluated for correctness, the resulting code is likely to be shorter and clearer, and the architecture is well suited to distributed, concurrent, and on demand tasks\cite{huHowFunctionalProgramming2015}.

@@

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/intro/continuity.png}
    \caption{The bar chart (b) and line plot (c) are different visual encodings of the precipitation data in the table (a). The bar chart preserves the \textit{continuity} of the stations by encoding the discrete station data as discrete bars (b). In contrast, the line plot (c) does not preserve this discrete \textit{continuity} because it connects the total temperature at each station by drawing a line through each point. Doing so implies that the total temperatures are points along a 1D continuous line, whereas there is no connectivity between the stations or their corresponding rows in the table.}
    \label{fig:intro:continuity}
\end{figure}
The notion that data is equivalent to visual representations when structure is preserved serves as the basis for visualization best practices. When \textit{continuity} is preserved, as in the bar chart (b) in \autoref{fig:intro:continuity}, then the graphic has not introduced new structure into the data.  In \autoref{fig:intro:continuity}, the line plot (c) does not preserve \textit{continuity} because the line connecting the total precipitation of the stations implies that the stations are connected to each other along a 1D continuous line. Bertin asserted that \textit{continuity} was preserved by choosing graphical marks that match the \textit{continuity} of the data - for example discrete data is a point, 1D continuous is the line, and 2D data is the area mark. Informally, Norman's Naturalness Principal\cite{NaturalnessPrincipleInfoVis} states that a visualization is easier to understand when the properties of the visualization match the properties of the data. This principal is made more concrete in Tufte's concept of graphical integrity, which is that a visual representation of quantitative data must be directly proportional to the numerical quantities it represents (Lie Principal), must have the same number of visual dimensions as the data, and should be well labeled and contextualized, and not have any extraneous visual elements\cite{tufteVisualDisplayQuantitative2001}. Expressivity, as defined by Mackinlay, is a measure how much of the mathematical structure in the data can be expressed in the visualizations; for example that ordered variables can be mapped into ordered visual elements. We generalize these different codifications of structure preserving encodings, proposing that a graphic is an equivalent representation of the data when \textit{continuity} and \textit{equivariance} are preserved.

@@

\begin{enumerate}
  \item formalization of the topology preserving relationship between data and graphic via continuous maps \autoref{sec:math:graphic:base}
  \item formalization of property preservation from data component to visual representation as equivariant maps \autoref{sec:math:artist:nu}
  \item functional oriented visualization architecture built on the mathematical model to demonstrate the utility of the model \autoref{sec:math:artist:q}
  \item prototype of the architecture built on Matplotlib's infrastructure to demonstrate the feasibility of the model. \autoref{sec:code:artists}
\end{enumerate}
We validate our model by using it to re-design artist and data access layer of Matplotlib, a general purpose visualization tool. We evaluate whether the redesign is successful by comparing it to existing implementation, recreating existing domain specific functionality with the new components, and by implementing components that provide new functionality in Matplotlib. While much of this functionality is currently possible in Matplotlib, a functional approach allows us to implement components in a more robust, modular, reliable way.
\end{document}

>>>>

Visualizations are representations of data and therefore reflect something of the underlying structure and semantics\cite{friendlyBriefHistoryData2008}, whether through direct mappings from data into visual elements or via figurative representations that have meaning due to their similarity in shape to external concepts \cite{byrneAcquiredCodesMeaning2016}. We define visualization components to be structure preserving maps from data to visual representations. As illustrated in \autoref{fig:intro:artist}, visualizations map weather station precipitation data (a) into graphical elements such as points (b) and lines (c) and map individual data components (columns) into components of the visualization such as position or color. For example, the gray position encoder function converts the latitude and longitude to x and y positions, and the color encoders map temperature and station name to colors. A \texttt{point} function composites these encodings into attributes of a point in the map (a). The \texttt{line} function composites encoders that convert station name to color and date and precipitation to x and y positions into a piece of a line in the time series plot (c). We identify the structure these maps must preserve as the \textit{continuity} of the data and the \textit{equivariance} of the data and visual components.

We introduce a model of visualization components based on these \textit{continuity} and \textit{equivariance} constraints and use this model to develop a design specification. Our specification is targeted at building block libraries, which Wongsuphasawat defines as visualization tools that provide independent functions that map components of data to visual representations\cite{wongsuphasawatNavigatingWideWorld2021}.  Just as the number of ways to arrange physical building blocks is solely constrained by the size and shape of the blocks, we propose that the only restriction on how building block library components can be composed are that the compositions preserve \textit{continuity} and \textit{equivariance}. Independent, modular, components are inherently functional\cite{hughesWhyFunctionalProgramming1989}, so we propose a functional architecture for our model. Doing so means the functional architecture can be evaluated for correctness, the resulting code is likely to be shorter and clearer, and the architecture is well suited to distributed, concurrent, and on demand tasks\cite{huHowFunctionalProgramming2015}.

@@

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/intro/continuity.png}
    \caption{The bar chart (b) and line plot (c) are different visual encodings of the precipitation data in the table (a). The bar chart preserves the \textit{continuity} of the stations by encoding the discrete station data as discrete bars (b). In contrast, the line plot (c) does not preserve this discrete \textit{continuity} because it connects the total temperature at each station with a line. Doing so implies that the total temperatures are points along a 1D continuous line, whereas there is no connectivity between the stations or their corresponding rows in the table.}
    \label{fig:intro:continuity}
\end{figure}
The notion that data is equivalent to visual representations when structure is preserved serves as the basis for visualization best practices. When \textit{continuity} is preserved, as in the bar chart (b) in \autoref{fig:intro:continuity}, then the graphic has not introduced new structure into the data.  In \autoref{fig:intro:continuity}, the line plot (c) does not preserve \textit{continuity} because the line connecting the total precipitation of the stations implies that the stations are connected to each other along a 1D continuous line. Bertin asserted that \textit{continuity} was preserved by choosing graphical marks that match the \textit{continuity} of the data - for example discrete data is a point, 1D continuous is the line, and 2D data is the area mark. Informally, Norman's Naturalness Principal\cite{NaturalnessPrincipleInfoVis,norman_things_smart} states that a visualization is easier to understand when the properties of the visualization match the properties of the data. This principal is made more concrete in Tufte's concept of graphical integrity, which is that a visual representation of quantitative data must be directly proportional to the numerical quantities it represents (Lie Principal), must have the same number of visual dimensions as the data, and should be well labeled and contextualized, and not have any extraneous visual elements\cite{tufteVisualDisplayQuantitative2001}. Expressivity, as defined by Mackinlay, is a measure how much of the mathematical structure in the data can be expressed in the visualizations; for example that ordered variables can be mapped into ordered visual elements. We generalize these different codifications of structure preserving encodings, proposing that a graphic is an equivalent representation of the data when \textit{continuity} and \textit{equivariance} are preserved.

@@

\begin{enumerate}
  \item formalization of the topology preserving relationship between data and graphic via continuous maps \autoref{sec:math:graphic:base}
  \item formalization of property preservation from data component to visual representation as equivariant maps \autoref{sec:math:artist:nu}
  \item functional oriented visualization architecture built on the mathematical model to demonstrate the utility of the model \autoref{sec:math:artist:q}
  \item prototype of the architecture built on Matplotlib's infrastructure to demonstrate the feasibility of the model. \autoref{sec:code:artists}
\end{enumerate}
We validate our model by using it to re-design the artist and data access layer of Matplotlib. We evaluate whether the redesign is successful by comparing it to existing implementation, recreating existing domain specific functionality with the new components, and by implementing components that provide new functionality in Matplotlib. While much of this functionality is currently possible in Matplotlib, a functional approach allows us to implement components in a more robust, modular, reliable way.
\end{document}
